---
title: "dplyr"
author: "Sebastian Kaltenbach"
output: html_notebook
---

```{r}
library(dplyr)
```

## distinct() Function in Dplyr  –  Remove duplicate rows of a dataframe:

```{r}
mydata <- mtcars

# Remove duplicate rows of the dataframe
distinct(mydata)
```

## Drop by column names in Dplyr:

select() function along with minus which is used to drop the columns by name
```{r}
mydata <- mtcars

# Drop the columns of the dataframe
select (mydata,-c(mpg,cyl,wt))
```

## Drop 3rd, 4th and 5th columns of the dataframe:

```{r}
mydata <- mtcars

# Drop 3rd,4th and 5th columns of the dataframe

select(mydata,-c(3,4,5))
```

## Re arrange or Re order the column of dataframe in R using Dplyr

```{r}
mydata <- mtcars

# Reorder the columns of the dataframe

Mydata1 = select(mydata, gear,hp,qsec,vs, everything())
Mydata1
```

## Rename the column name in R using Dplyr

```{r}
mydata <- mtcars

# Rename the column name of the dataframe
Mydata1 = rename(mydata, displacement=disp, cylinder=cyl)
Mydata1
```

## Filter or subsetting the rows in R using Dplyr

```{r}
mydata <- mtcars

# subset the rows of dataframe with condition
Mydata1 = filter(mydata,cyl==6)
Mydata1
```

## Filter or subsetting the rows in R with multiple conditions using Dplyr

```{r}
mydata <- mtcars

# subset the rows of dataframe with multiple conditions
Mydata1 = filter(mydata, gear %in% c(4,5))
Mydata1
```

## Summary of column in dataset in R using Dplyr – summarise()

```{r}
mydata <- mtcars

# summarise the columns of dataframe
summarise(mydata, mpg_mean=mean(mpg),mpg_median=median(mpg))
```

## Create new variable in R using Mutate Function in dplyr

```{r}
mydata2 <- iris 

# Mutate function for creating new variable to the dataframe in R

mydata3 = mutate(mydata2, sepal_length_width_ratio=Sepal.Length/Sepal.Width)
head(mydata3)
```

## Join dataframes

```{r}
df_primary <- tribble(
  ~ID, ~y,
   "A", 5,
   "B", 5,
   "C", 8,
   "D", 0,
  "F", 9)
df_secondary <- tribble(
  ~ID, ~z,
   "A", 30,
   "B", 21,
   "C", 22,
   "D", 25,
   "E", 29)
```

### Dplyr left_join()

The most common way to merge two datasets is to use the left_join() function. We can see from the picture below that the key-pair matches perfectly the rows A, B, C and D from both datasets. However, E and F are left over. How do we treat these two observations? With the left_join(), we will keep all the variables in the original table and don’t consider the variables that do not have a key-paired in the destination table. In our example, the variable E does not exist in table 1. Therefore, the row will be dropped. The variable F comes from the origin table; it will be kept after the left_join() and return NA in the column z. The figure below reproduces what will happen with a left_join().
```{r}
left_join(df_primary, df_secondary, by ='ID')
```

### Dplyr right_join()

The right_join() function works exactly like left_join(). The only difference is the row dropped. The value E, available in the destination data frame, exists in the new table and takes the value NA for the column y.
```{r}
right_join(df_primary, df_secondary, by = 'ID')
```

### Dplyr inner_join()

When we are 100% sure that the two datasets won’t match, we can consider to return only rows existing in both dataset. This is possible when we need a clean dataset or when we don’t want to impute missing values with the mean or median.
The inner_join()comes to help. This function excludes the unmatched rows.
```{r}
inner_join(df_primary, df_secondary, by ='ID')
```

### Dplyr full_join()

Finally, the full_join() function keeps all observations and replace missing values with NA.
```{r}
full_join(df_primary, df_secondary, by = 'ID')
```

### Dplyr slice()

Create an exemplary data frame to visualize functionality of slice-function.
```{r}
data <- data.frame(x1 = 1:5,                    
                   x2 = LETTERS[1:5],
                   x3 = 5)
data 
```

The slice function of the dplyr R package can be used to extract columns of a data frame or tibble as shown below:
```{r}
slice(data, c(1, 3, 5)) 
```